```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
if (!require("pacman")){
    install.packages("pacman")
} #if pacman is not installed, install it
pacman::p_load(tidyverse, data.table, ggplot2,janitor,ggpubr,ggrepel,
               devtools,readxl) #pacman will now install and load these packages if you don't have them

if (!require("annotables")) {
    devtools::install_github("stephenturner/annotables")
}

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("PCAtools")
library(PCAtools)
```
This is the vignette for the PCAtools package
https://bioconductor.org/packages/release/bioc/vignettes/PCAtools/inst/doc/PCAtools.html

We're going to put the "ensgene" into a 'rowname' and then do the PCA analysis.
Now I'm going to write this using the pipe operator "%>%". Your "hot key" to type that in is: Ctrl + Shift + M

```{r}
bdnf_counts <- featureCounts[,1:9] # so that you exclude the "gene name" column
colnames(bdnf_counts) <- c("ensgene", "CTRL_chx_1", "CTRL_chx_2", "CTRL_chx_3", "CTRL_chx_4",
                           "DOX_chx_1", "DOX_chx_2", "DOX_chx_3", "DOX_chx_4")

pca <- bdnf_counts %>% 
    column_to_rownames('ensgene') #%>% t()
pca_res <- prcomp(t(pca),scale = F)
```

So let's do a PCA plot using meta data. You'll find formatted meta data in the drive

```{r}
path_to_meta_you_downloaded <-  "/Users/matteozanovello/Documents/phd/research_lines/splicing_chx/data/metadata.csv" #update this
meta_df <- fread(path_to_meta_you_downloaded)
meta_df <- meta_df %>% 
    column_to_rownames('sample') 

p <- PCAtools::pca(pca, metadata = meta_df, removeVar = 0.1)
```
What does the removeVar argument do? Why did I include it? 

As before the screeplot
```{r}
screeplot(p, axisLabSize = 18, titleLabSize = 22)
```

Now we're going to make what is called an "eigencorplot"

This is going to correlate our metadata variables with our principle components. 
Star means "significant". What kind of correlation is it? Check out the help documentation 
by entering `?PCAtools::eigencorplot` in the console
```{r}
PCAtools::eigencorplot(p,components = getComponents(p),
                       metavars = c('condition',"treatment"))
```

We can do a biplot with the 2 PC's that are correlating

```{r}
biplot(p,
       x = "PC1",
       y = "PC2",
       colby = 'condition')
```

Let's go ahead and find out what the genes are that are correlated with PC1.
Let's directly take the loadings of each of the genes (a PCA is a linear combination). 

I'm going to also show you now how to put the symbol back on the table. 
We shall use the wonder of the `left_join` 

```{r}
pc1_gene_loadings = p$loadings %>% 
    dplyr::select(PC1) %>% 
    rownames_to_column('ensgene') %>% 
    left_join(annotables::grch38 %>% dplyr::select(ensgene)) %>% 
#https://statisticsglobe.com/r-dplyr-join-inner-left-right-full-semi-anti
    arrange(-PC1)
```

Now we've seen a join let's do the same thing with our original count table
```{r}
bdnf_counter = bdnf_counts %>% 
    left_join(annotables::grch38 %>% dplyr::select(ensgene))
```


I want to look at the expression of the top 15 genes on PC1 so I'm going 
to plot their expression.

```{r}
for_ggplot <- bdnf_counter %>% 
    left_join(pc1_gene_loadings %>% head(15)) %>%
    filter(!is.na(PC1)) %>%
    #dplyr::select(-ensgene) %>% 
    melt(id.vars = c("ensgene","PC1")) %>%
    mutate(ensgene = fct_reorder(ensgene,-PC1)) %>% 
    separate(variable, into = c("condition","treatment"))
ggplot(for_ggplot, aes(x = treatment, y = value, fill = condition)) + 
    geom_col(position = "dodge2") + facet_wrap(~ensgene)

```

4su labelling is overwhelming the pc1 so i do analysis without those 2 timepoints, removing BDNF_6h, CONTROL_6h

```{r}
bdnf_counts <- select(bdnf_counts, -c(BDNF_6h, CONTROL_6h))

meta_df <- meta_df %>% filter(labelling_time != 6)

bdnf_counts <- select(bdnf_counts, -(symbol)) %>% 
    unique() %>% 
    remove_rownames() %>% 
    column_to_rownames("ensgene")
```

what genes account for PC3 w/o BDNF6h, Control6h

```{r}
pc3_gene_loadings = p$loadings %>% 
    dplyr::select(PC3) %>% 
    rownames_to_column('ensgene') %>% 
    left_join(annotables::grch38 %>% dplyr::select(ensgene,symbol)) %>% 
#https://statisticsglobe.com/r-dplyr-join-inner-left-right-full-semi-anti
    arrange(-PC3)

bdnf_counts %>% 
    rownames_to_column("ensgene") %>% 
    left_join(pc3_gene_loadings %>% head(15))  %>% 
    filter(!is.na(PC3)) %>% 
    select(-ensgene) %>% 
    melt(id.vars = c("symbol","PC3")) %>% 
    mutate(symbol = fct_reorder(symbol,-PC3)) %>% 
    separate(variable, into = c("condition","BDNF")) %>%
    ggplot(aes(x = BDNF, y = value, fill = condition)) +
    geom_col(position = "dodge2") + 
    facet_wrap(~symbol)
```

