---
title: "CHX dependent analysis"
author: "Matteo Zanovello"
date: "2022-01-10"
---

```{r setup, include = False}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
if (!require("pacman")){
    install.packages("pacman")
} #if pacman is not installed, install it
pacman::p_load(tidyverse, data.table, ggplot2, janitor, ggpubr, ggrepel,
               devtools, readxl, DESeq2, readr, edgeR, annotables)
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
library(PCAtools)
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)

source(here::here('scripts','feature_count','create_feature_count_table.R'))
source(here::here('scripts','feature_count','make_deseq_dfs.R'))
source(here::here('scripts','feature_count','run_standard_deseq.R'))
source(here::here('scripts','feature_count','make_volcano_plot.R'))
source(here::here('scripts','splicing_volcanoplot.R'))
```


```{r}
featureCounts <- create_feature_count_table("/Users/matteozanovello/Documents/phd/research_lines/splicing_chx/data/feature_count") #change here!!
#my_df <- make_deseq_dfs(featureCounts, grep_pattern = "", base_grep = "CTRL", contrast_grep = "DOX")

grepperz <- c("CTRL", "DOX")
for (k in grepperz) {
my.dds <- run_standard_deseq("/Users/matteozanovello/Documents/phd/research_lines/splicing_chx/data/feature_count", 
                             base_grep = "ctrl",
                             contrast_grep = "chx",  
                             grep_pattern = k,
                             baseName = "Untreated",
                             contrastName = 'Cycloheximide')

ppp <- label_significant(my.dds$results_table, log2FoldCut = 2.5, log10padj = 40) + annotate("text", x=8, y=-5, label=k)
plot(ppp)
}
```

```{r}
grepper <- c("ctrl", "_chx")
for (j in grepper) {
my.dds <- run_standard_deseq("/Users/matteozanovello/Documents/phd/research_lines/splicing_chx/data/feature_count", 
                             base_grep = "CTRL",
                             contrast_grep = "DOX",  
                             grep_pattern = j,
                             baseName = "Untreated",
                             contrastName = 'TDP43KD')

bbb <- label_significant(my.dds$results_table, log2FoldCut = 2, log10padj = 15) + annotate("text", x=7, y=-0.5, label=j)
plot(bbb)
}
```

splicing volcano
```{r}
input <- c(#"ControlControl-ControlTDP43KD") #, "ControlControl-CycloheximideControl", "ControlTDP43KD-CycloheximideTDP43KD",
    "CycloheximideControl-CycloheximideTDP43KD")
for (i in input) {
    input_csv <- paste(i, "csv", sep = ".")
input_splicing <- fread(file.path(here::here(), "data", input_csv))
input_splicing <- separate_rows(input_splicing, c(names(input_splicing[,c(4:9,12,15)])), sep  = c(";"), convert = T)

a <- splicing_dots_tables_function(input_splicing) #+ annotate("text", x=0.7, y=0.1, label=i)
plot(a)
}
```

This is the vignette for the PCAtools package
https://bioconductor.org/packages/release/bioc/vignettes/PCAtools/inst/doc/PCAtools.html

We're going to put the "ensgene" into a 'rowname' and then do the PCA analysis using meta data
Now I'm going to write this using the pipe operator "%>%". Your "hot key" to type that in is: Ctrl + Shift + M
```{r}
bdnf_counts <- featureCounts
colnames(bdnf_counts) <- c("ensgene", 
                           "CTRL_chx_1", "CTRL_chx_2", "CTRL_chx_3", "CTRL_chx_4",
                           "CTRL_ctrl_1", "CTRL_ctrl_2", "CTRL_ctrl_3", "CTRL_ctrl_4",
                           "DOX_chx_1", "DOX_chx_2", "DOX_chx_3", "DOX_chx_4", 
                           "DOX_ctrl_1", "DOX_ctrl_2", "DOX_ctrl_3", "DOX_ctrl_4", 
                           "symbol")

pca_table <- bdnf_counts[,1:17] %>% 
    column_to_rownames('ensgene')

path_to_meta_you_downloaded <-  "/Users/matteozanovello/Documents/phd/research_lines/splicing_chx/data/metadata.csv" #update this
meta_df <- fread(path_to_meta_you_downloaded)
meta_df <- meta_df %>% 
    column_to_rownames('sample')

p <- PCAtools::pca(pca_table, metadata = meta_df, removeVar = 0.1)
screeplot(p, axisLabSize = 18, titleLabSize = 22)
```

Now we're going to make what is called an "eigencorplot"
This is going to correlate our metadata variables with our principle components. 
```{r}
PCAtools::eigencorplot(p, components = getComponents(p),
                       metavars = c('condition',"treatment"))
```

We can do a biplot with the 2 PC's that are correlating
```{r}
biplot(p,
       x = "PC1", #0.99 with treatment
       y = "PC2", #0.97 with condition
       colby = c('condition'),
       axisLabSize = 24,
       labSize = 6)
```

Let's go ahead and find out what the genes are that are correlated with PC2.
Let's directly take the loadings of each of the genes (a PCA is a linear combination). 
I'm going to also show you now how to put the symbol back on the table. 
We shall use the wonder of the `left_join`
```{r}
pc1_gene_loadings = p$loadings %>% 
    dplyr::select(PC2) %>% 
    rownames_to_column('ensgene') %>% 
    left_join(annotables::grch38 %>% dplyr::select(c(ensgene))) %>% 
#https://statisticsglobe.com/r-dplyr-join-inner-left-right-full-semi-anti
    arrange(-PC2)

bdnf_counts %>% 
    left_join(annotables::grch38 %>% dplyr::select(c(ensgene,symbol))) %>% 
    left_join(pc1_gene_loadings %>% head(16)) %>%
    filter(!is.na(PC2)) %>%
    dplyr::select(-ensgene) %>% 
    melt(id.vars = c("symbol","PC2")) %>%
    mutate(ensgene = fct_reorder(symbol,-PC2)) %>% 
    separate(variable, into = c("condition","treatment")) %>%
    ggplot(aes(x = condition, y = value, fill = treatment)) + 
    geom_col(position = "dodge2") + facet_wrap(~symbol)

```

GO analysis https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-go.html
#see notes above
```{r}
results <- my.dds$results_table
filtered_res <- filter(results, padj < 0.1 & abs(log2FoldChange) > 2)
## feature 1: numeric vector
geneList <- pull(filtered_res,3)
## feature 2: named vector
names(geneList) = as.character(pull(filtered_res,8))
## feature 3: decreasing order
geneList = sort(geneList, decreasing = TRUE)

ggo <- groupGO(gene = names(geneList),
               OrgDb = org.Hs.eg.db,
               keyType = "SYMBOL",
               ont = "BP",
               level = 3)
cnetplot(ggo)
```

```{r}
ego <- enrichGO(gene = names(geneList),
                keyType = "SYMBOL",
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff = 1,
                qvalueCutoff = 1)
goplot(ego)
dotplot(ego, showCategory=30)
cnetplot(ego, foldChange=geneList)
```

```{r}
pgo <- pairwise_termsim(ego)
emapplot(pgo)
```

```{r}
gse <- gseGO(geneList     = geneList,
              keyType = "SYMBOL",
              OrgDb        = org.Hs.eg.db,
              ont          = "CC",
              minGSSize    = 50,
              maxGSSize    = 200,
              pvalueCutoff = 0.01,
              verbose      = FALSE)
gseaplot(gse, geneSetID="GO:0043231")
```